Wrangle transcripts

### 2025-10-28 code/demog.qmd
clean up of transcripts for referential, matching ids

```{r}
expt_1 <- read_csv(here("data/expt_1/clean_data.csv")) |>
  select(game, speaker, listener) |>
  pivot_longer(c("speaker", "listener"), values_to = "id") |>
  select(-name) |>
  unique() |>
  mutate(expt = 1) |>
  mutate(sid = id)

id_sub <- tribble(
  ~id, ~sid,
  "apple123", "id74",
  "pear124", "id75",
  "apple127", "id76",
  "pear128", "id77",
  "80", "id78",
  "81", "id79",
  "apple111", "id90",
  "pear222", "id91",
  "apple333", "id92",
  "pear444", "id93",
  "apple255", "id94",
  "pear257", "id95"
)

expt_2 <- read_csv(here("data/expt_2/clean_data.csv")) |>
  select(game = gameId, speaker, listener) |>
  pivot_longer(c("speaker", "listener"), values_to = "id") |>
  select(-name) |>
  unique() |>
  mutate(expt = 2) |>
  left_join(id_sub) |>
  mutate(sid = case_when(
    !is.na(sid) ~ sid,
    str_detect(id, "id") ~ id,
    T ~ str_c("id", id)
  ))

@@

```{r}
demogs <- raw_demog |>
  mutate(
    date_test = mdy(date_test),
    date_birth = mdy(date_birth)
  ) |>
  select(sid, included, date_test, date_birth, hispanic, ethnicity, gender) |>
  full_join(kids) |>
  filter(!is.na(included)) |>
  mutate(
    age_days = (date_test - date_birth) |> as.numeric(),
    age_month = floor(age_days / 30.43)
  ) |>
  select(-age_days) |>
  select(expt, id, game, sid, gender, age_month, ethnicity, hispanic) |>
  write_csv(here("data/clean_demog_1_2.csv"))
```

>>>>

```{r}
expt_1 <- read_csv(here("data/expt_1/clean_data.csv")) |>
  select(game, speaker, listener) |>
  pivot_longer(c("speaker", "listener"), values_to = "id") |>
  select(-name) |>
  unique() |>
  mutate(expt = 1)

expt_2 <- read_csv(here("data/expt_2/clean_data.csv")) |>
  select(game, speaker, listener) |>
  pivot_longer(c("speaker", "listener"), values_to = "id") |>
  select(-name) |>
  unique() |>
  mutate(expt = 2)

@@

```{r}
demogs <- raw_demog |>
  mutate(
    date_test = mdy(date_test),
    date_birth = mdy(date_birth)
  ) |>
  select(id=sid, included, date_test, date_birth, hispanic, ethnicity, gender) |> 
  full_join(kids) |>
  filter(!is.na(included)) |>
  mutate(
    age_days = (date_test - date_birth) |> as.numeric(),
    age_month = floor(age_days / 30.43)
  ) |>
  select(-age_days) |>
  mutate(gender=str_to_lower(gender)) |> 
  select(expt, id, game, gender, age_month, ethnicity, hispanic)

demogs |> filter(expt==1) |> write_csv(here("data/expt_1/clean_demog.csv"))

demogs |> filter(expt==2) |> write_csv(here("data/expt_2/clean_demog.csv"))

```

# expt 3

```{r}

expt_3_demog <- read_csv(here("expt_3_demog.csv"))

expt_3 <- read_csv(here("data/expt_3/clean_data.csv")) |>
  select(game, speaker, listener) |>
  pivot_longer(c("speaker", "listener"), values_to = "id") |>
  select(-name) |>
  unique() |> 
  mutate(expt=3)

expt_3_demog <- read_csv(here("expt_3_demog.csv")) |> 
 mutate(
    date_test = mdy(date_test),
    date_birth = ymd(date_birth),
    id=str_c("id", SID)
  ) |>
  select(id, date_test, date_birth, hispanic, ethnicity, gender) |> 
  full_join(expt_3) |> 
  filter(!is.na(game)) |> 
  mutate(
    age_days = (date_test - date_birth) |> as.numeric(),
    age_month = floor(age_days / 30.43)
  ) |>
  select(-age_days) |>
  mutate(gender=str_to_lower(gender)) |> 
  select(expt, game, id, gender, age_month, ethnicity, hispanic) |> write_csv(here("data/expt_3/clean_demog.csv"))

```

### 2025-10-28 code/expt_1/analysis_1.Rmd
clean up of transcripts for referential, matching ids

knitr::opts_chunk$set(
                      echo=F, warning=F, cache=T, 
                      message=F, sanitize = T)
theme_set(theme_classic())
transcript <- read_csv(here("data/timed_transcript.csv"))

@@

@@

Trials were excluded if the speaker did not provide any description before the listener clicked.

>>>>

knitr::opts_chunk$set(
                      echo=F, warning=F, cache=T, 
                      message=F, sanitize = T)
theme_set(theme_classic())

transcript <- read_csv(here("data/timed_transcript.csv"))

@@

```
.

# Exclusions

```{r}
transcript <- read_csv(here(data_loc, "transcripts.csv"))
data <- read_csv(here(data_loc, "clean_data.csv"))
link <- read_tsv(here(data_loc, "link_transcripts.csv"))

data <- read_csv(here(data_loc, "clean_data.csv")) |>
  rename(trial = trialNum) |>
  inner_join(link |> filter(is.na(exclude)))

transcript <- transcript |>
    inner_join(link |> filter(is.na(exclude))) |> 
  filter(!is.na(description))

words <- transcript |> mutate(words=description |> str_count("\\S+")) |>
  filter(role=="S") |>
  group_by(gameId, game, trial) |>
  summarize(words=sum(words))

all_data <- data |>left_join(words) |> 
  filter(!is.na(words)) |> 
   mutate(type = case_when(
    trial < 4 ~ "practice",
    trial < 8 ~ "block 1",
    trial < 12 ~ "block 2",
    trial < 16 ~ "block 3",
    trial < 20 ~ "block 4"
  )) |>
  filter(!is.na(response)) |> 
  mutate(correct=as.numeric(correct))

all_data |> select(game, speaker) |> distinct()

@@

### 2025-10-28 code/expt_1/data_prep.Rmd
clean up of transcripts for referential, matching ids

pilot_start=lubridate::ymd('2023-07-07') 
pilot_end=lubridate::ymd('2023-07-12')

real_start=lubridate::ymd('2023-07-17')
real_end=lubridate::ymd('2023-08-09')

@@

pilot_ids <- str_c("id", 3:12) # Ids 1, 2 were adult pilot test

real_ids <-str_c("id", c(13:16, 19:36, 39:54, 57:58)) # skipping kids who didn't complete enough trials
```

@@

players <- read_csv(here(player_source)) |> filter(id %in% real_ids) %>% select(id, playerId=`_id`)

games <- read_csv(here("data/expt_1/games.csv"), col_types="ccc")|> filter(is.na(exclude)) |>  mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)

@@

raw_data <- read.csv(here(data_source), header = TRUE) %>% 
  filter(createdAt >= real_start) %>%
  filter(createdAt <= real_end) %>% 
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  select(-chat) |> 
  pivot_longer(cols=starts_with('player'), names_to=c("playerId", "info"), names_prefix="player_", names_sep="_",
               values_to="values", values_transform=as.character, values_drop_na=T) %>% 
  filter(values!="") %>% 
  filter(playerId!=speaker) %>% 
  pivot_wider(names_from=info, values_from=values) |> 
  left_join(players, by=c("playerId")) |> 
  select(-playerId) |> 
  rename(playerId=id) |> 
  left_join(players, by=c("speaker"="playerId")) |> 
  select(-speaker) |> 
  rename(speaker=id) %>% 
  left_join(games) |> 
  rename(game=gameConfig) |> 
  select(-gameId, -offset_sec) |> 
  filter(!is.na(game)) |> 
  mutate(time=as.numeric(time))

@@

less_raw_data <- raw_data %>% select(game, target, distractor, trialNum, targetNum, repNum, response, correct, time, listener=playerId, speaker) %>% 
  mutate(across(c(target, distractor, response), ~ label_image(.x))) %>% write_csv(here("data/clean_data.csv"))
```

```{r}
#just for pilot, need to fix offset
#fix <- raw_data |> filter(trialNum==0) |> select(game, time)

#real_offset <- games |> left_join(fix, by=c("gameConfig"="game")) |> mutate(offset_sec=offset_sec-time/1000)

real_offset <- games

@@

# Transcript wrangle - new

wrangle_transcript <- function(game_num){
 read_csv(here("data/expt_1/clean_transcripts", str_c(game_num,".csv")), show_col_types=F) |> 
  select(start, end, text, speaker, trial, role, echo, description, backchannel) |> 
    fill(trial) %>% 
  mutate(start=start/1000, end=end/1000) |> 
    nest(data=c(start, end, text, speaker, trial, role, echo, description, backchannel))
}

#wrangle_transcript("game41") %>% unnest(data) %>% View()

real_offset$transcripts <- map_df(real_offset$gameConfig, wrangle_transcript)

transcripts <- real_offset |> unnest(transcripts) |> unnest(data)

# this section is janky as hell, but it does work

stages <- read_csv(here("data/expt_1/stages.csv")) |> select(index, gameId, startTimeAt) |> group_by(gameId)

gameStarts <- stages |> filter(index==0) |> rename(first=startTimeAt) |> select(-index)

stage_timing <- stages |> left_join(gameStarts) |>
  mutate(rel_time=startTimeAt-first) |> 
  mutate(rel_time=lubridate::as.period(rel_time) |> lubridate::period_to_seconds()) |> 
   mutate(trialNum=index/2,
          start=rel_time,
          end=lead(rel_time)) |> 
  select(gameId, trialNum,start, end) |>  left_join(games) |>  ungroup() |> 
  mutate(trialNumStart=trialNum, trialNumEnd=trialNum) |> 
  select(game=gameConfig, trialNumStart, trialNumEnd, start, end) |> filter(!is.na(game)) |> filter(!is.na(start)) |> filter(!is.na(end))
```

```{r}
timed_transcripts <-  transcripts |>
  mutate(start=start-offset_sec, end=end-offset_sec) |> 
  select(-offset_sec, -gameId) |> 
  mutate(trialNumStart=NA, trialNumEnd=NA) |> 
  rename(game=gameConfig) |> 
  bind_rows(stage_timing) |> 
  group_by(game) |> 
  arrange(start) |> 
  fill(trialNumStart) |> 
  arrange(end) |> 
  fill(trialNumEnd, .direction="up") |> 
  arrange(game) |> 
  filter(!is.na(text))

echoes <- timed_transcripts %>% select(game,trial, echo) %>% filter(echo=="Y") %>% unique()

trial_timing <- stage_timing %>% select(game, trial=trialNumStart, trial_start=start)

transcript_full <- timed_transcripts %>% select(-echo, -exclude) %>% left_join(echoes) %>% 
  left_join(trial_timing) %>% 
  rename(start_time=start, end_time=end) %>% 
  write_csv(here("data/expt_1/timed_transcript.csv"))
```

# Checks
```{r}
# Quick check -- for kid talking, so auto-trial and hand-trial align?

transcript_full %>% filter(!is.na(trialNumStart)) %>% filter(str_detect(speaker, "id")) %>% filter(!is.na(role)) %>% filter(trial!=trialNumStart)

# we'll go with the by-hand method -- these mostly seem to be when there's a wide span of time! possibly due to whisper being silly or us hand filling due to whisper missing things

checks on data:
* anything that has a description should be from a speaker and have Y for backchannel
* anything that has a Y for backchannel should be from a child

```{r}
transcript_full %>% filter(!is.na(description)) %>% filter(is.na(backchannel))
transcript_full %>% filter(!is.na(description)) %>% ungroup() %>% select(role) %>% unique()
transcript_full %>% filter(backchannel=="Y") %>% ungroup() %>% select(role) %>% unique() 
transcript_full %>% filter(backchannel=="Y") %>% ungroup() %>% filter(!str_detect(speaker,"id"))
```

@@

```{r}
transcript <- read_csv(here("data/expt_1/timed_transcript.csv"))

dat <- read_csv(here("data/expt_1/clean_data.csv")) %>% select(trial=trialNum, game, target, distractor)

transcript %>% filter(!is.na(description)) %>% group_by(trial, game) %>% 
  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% left_join(dat) %>% write_csv(here("data/expt_1/pre_sbert.csv"))

@@

similarities <- M_mat %>%
    mutate(combinedId=str_c(game,trial,sep="_")) %>%
    make_across_df(F_mat, 'cosine') %>%
    separate(dim1, into=c("game1","trial1"), convert=T, sep="_") %>%
    separate(dim2, into=c("game2","trial2"), convert=T, sep="_") %>%
    mutate(sim = ifelse(is.nan(sim), NA, sim)) %>% write_rds(here("data/expt_1/similarities.rds"))

>>>>

@@

@@

players <- read_csv(here(player_source)) |> select(id, playerId=`_id`)

@@

raw_data <- read.csv(here(data_source), header = TRUE) %>% 
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  select(-chat) |> 
  pivot_longer(cols=starts_with('player'), names_to=c("playerId", "info"), names_prefix="player_", names_sep="_",
               values_to="values", values_transform=as.character, values_drop_na=T) %>% 
  filter(values!="") %>% 
  filter(playerId!=speaker) %>% 
  pivot_wider(names_from=info, values_from=values) |> 
  left_join(players, by=c("playerId")) |> 
  select(-playerId) |> 
  rename(playerId=id) |> 
  left_join(players, by=c("speaker"="playerId")) |> 
  select(-speaker) |> 
  rename(speaker=id)

@@

less_raw_data <- raw_data %>% select(gameId, target, distractor, trialNum, targetNum, repNum, response, correct, time, listener=playerId, speaker) %>% 
  mutate(across(c(target, distractor, response), ~ label_image(.x))) %>%
 inner_join(read_csv(here("data/expt_1/link_transcripts.csv")) |> select(gameId, game)) |> 
write_csv(here("data/expt_1/clean_data.csv"))

@@

# Transcript wrangle
```{r}
wrangle_transcript <- function(game_num) {
  if(file.exists(here("data/expt_1/raw_transcripts", str_c(game_num, ".csv")))){
  file = read_csv(here("data/expt_1/raw_transcripts", str_c(game_num, ".csv")))
  }
  else{
    if(file.exists(here("data/expt_1/raw_transcripts/exclude", str_c(game_num, ".csv")))){
        #file = read_csv(here("data/expt_1/raw_transcripts/exclude", str_c(game_num, ".csv")))
         file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
    }
    else{
      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
    }
  }
  file|>
    select(start, end, text, speaker, trial, role, echo, description) |>
    fill(trial) |> 
    mutate(start = start / 1000, end = end / 1000) |>
   nest(data = c(start, end, text, speaker, trial, role, echo, description))

}

transcripts <- read_csv(here("data/expt_1/link_transcripts.csv")) |> 
  select(gameId, game) |>
  mutate(transcript = map_df(game, wrangle_transcript)) |>
  unnest(transcript) |>
  unnest(data)

write_csv(transcripts, here("data/expt_1/transcripts.csv"))
```

@@

```{r}
transcript <- read_csv(here("data/expt_1/transcripts.csv"))

dat <- read_csv(here("data/expt_1/clean_data.csv")) %>% select(trial = trialNum, gameId, target, distractor)

transcript %>%
  filter(!is.na(description)) %>%
  group_by(trial, gameId) %>%
  mutate(trial = as.numeric(trial)) |>
  filter(trial > 1) |>
  summarize(words = str_c(description, sep = " ", collapse = ", ")) %>%
  left_join(dat) %>%
  write_csv(here("data/expt_1/pre_sbert.csv"))

@@

similarities <- M_mat %>%
    mutate(combinedId=str_c(gameId,trial,sep="_")) %>%
    make_across_df(F_mat, 'cosine') %>%
    separate(dim1, into=c("game1","trial1"), convert=T, sep="_") %>%
    separate(dim2, into=c("game2","trial2"), convert=T, sep="_") %>%
    mutate(sim = ifelse(is.nan(sim), NA, sim)) %>% write_rds(here("data/expt_1/similarities.rds"))

### 2025-10-28 code/expt_3/data_prep_3.Rmd
clean up of transcripts for referential, matching ids

real_ids <- c(str_c("", c(
  159:162, 167:170,
  173:174, 181:184,
  187:188, 191:196,
  199:200, 203:204,
  207:208, 227:234,
  237:240, 243:252
)))
# kids ids to include
```

@@

```{r}
players <- read_csv(here(player_source)) |>
  filter(id %in% real_ids) %>%
  select(id, playerId = `_id`)

# expect 50 kids (25 games)
```

```{r}
raw_data <- read.csv(here(data_source), header = TRUE) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  # select(-chat) |>
  pivot_longer(
    cols = starts_with("player"), names_to = c("playerId", "info"), names_prefix = "player_", names_sep = "_",
    values_to = "values", values_transform = as.character, values_drop_na = T
  ) %>%
  filter(values != "") %>%
  filter(playerId != speaker) %>%
  pivot_wider(names_from = info, values_from = values) |>
  left_join(players, by = c("playerId")) |>
  select(-playerId) |>
  rename(playerId = id) |>
  filter(!is.na(playerId)) |>
  left_join(players, by = c("speaker" = "playerId")) |>
  select(-speaker) |>
  rename(speaker = id)

@@

less_raw_data <- raw_data %>%
  select(target, distractor, trialNum, targetNum, repNum, response, correct, time, listener = playerId, speaker, gameId) %>%
  mutate(across(c(target, distractor, response), ~ label_image(.x))) %>%
  write_csv(here("data/expt_3/clean_data.csv"))
```

below here copied from expt 2 and not updated!

# Transcript wrangle - new

```{r}
library(readxl)

wrangle_transcript <- function(game_num) {
  read_tsv(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv"))) |>
    select(start, end, text, speaker, trial, role, echo, description) |>
    fill(trial) %>%
    mutate(start = start / 1000, end = end / 1000) |>
    mutate(trial = trial-1) |>
    nest(data = c(start, end, text, speaker, trial, role, echo, description))
}

transcripts <- read_tsv(here("data/expt_3/link_transcripts.csv")) |> 
  select(gameId, gameConfig) |>
  mutate(transcript = map_df(gameConfig, wrangle_transcript)) |>
  unnest(transcript) |>
  unnest(data)

@@

# Checks

how much echoing is there?

```{r}
transcripts |> filter(!is.na(echo))

# not much, yay!
```

@@

```{r}
library(reticulate)
np <- import("numpy")
mat <- np$load(here("data/expt_3/post_sbert.npy"))
saveRDS(mat, here("data/expt_3/post_sbert.RData"))
```

>>>>

```

@@

```{r}

id_sub <- function(id){
  new_id = case_when(
    id=="pear" ~ "id153",
    id=="apple" ~ "id154",
    id=="pear155" ~ "id155",
    id=="apple156" ~ "id156",
    id=="pear157" ~ "id157",
    id=="apple158" ~ "id158",
    id=="pear123" ~ "id163",
    id=="apple124" ~ "id164",
    id=="pear124"~"id165",
    id=="apple125" ~ "id166",
    T ~ str_c("id", id)
  )
  return(new_id)
}

players <- read_csv(here(player_source)) |> select(playerId=`_id`, id) |> rowwise() |>  mutate(id=id_sub(id)) |> ungroup()

raw_data <- read.csv(here(data_source), header = TRUE) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  pivot_longer(
    cols = starts_with("player"), names_to = c("playerId", "info"), names_prefix = "player_", names_sep = "_",
    values_to = "values", values_transform = as.character, values_drop_na = T
  ) %>%
  filter(values != "") %>%
  filter(playerId != speaker) %>%
  pivot_wider(names_from = info, values_from = values) |>
  left_join(players, by = c("playerId")) |>
  select(-playerId) |>
  rename(playerId = id) |>
  filter(!is.na(playerId)) |>
  left_join(players, by = c("speaker" = "playerId")) |>
  select(-speaker) |>
  rename(speaker = id)

@@

less_raw_data <- raw_data %>%
  select(target, distractor, trialNum, targetNum, repNum, response, correct, time, listener = playerId, speaker, gameId) %>%
  mutate(across(c(target, distractor, response), ~ label_image(.x))) |> 
  inner_join(read_tsv(here("data/expt_3/link_transcripts.csv")) |> select(gameId, game)) |> 
  write_csv(here("data/expt_3/clean_data.csv"))

```

# Transcript wrangle

```{r}

wrangle_transcript <- function(game_num) {
  if(file.exists(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv")))){
  file = read_tsv(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv")))
  }
  else{
    if(file.exists(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))){
        file = read_tsv(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))
    }
    else{
      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
    }
  }
  file|>
    select(start, end, text, speaker, trial, role, echo, description) |>
    fill(trial) %>%
    mutate(start = start / 1000, end = end / 1000) |>
    mutate(trial = trial-1) |>
    nest(data = c(start, end, text, speaker, trial, role, echo, description))

}

transcripts <- read_tsv(here("data/expt_3/link_transcripts.csv")) |> 
  select(gameId, game) |>
  mutate(transcript = map_df(game, wrangle_transcript)) |>
  unnest(transcript) |>
  unnest(data)

@@

@@

(go run sbert here)
```{r}
library(reticulate)
np <- import("numpy")
mat <- np$load(here("data/expt_3/post_sbert.npy"))
saveRDS(mat, here("data/expt_3/post_sbert.RData"))
```
